Effective Reionization Likelihood from Planck 2018 Data (?)

### 2021-03-22 paper/rei.tex

\begin{figure}[ht]
\includegraphics[width=0.45\textwidth]{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2.pdf}
\caption{Comparison of the 68\% and 95\% C.L. constraints on the cumulative optical depth with different Planck likelihoods. \textit{Top}: The Planck 2015 vs Planck 2018 likelihoods, the latter including the official low-$\ell$ $EE$ likelihood $\texttt{simallEE}$. \textit{Bottom}: The Planck 2018 likelihoods using $\texttt{simallEE}$ for the low-$\ell$ $EE$ likelihood vs \texttt{srollv2} which has better foreground cleaning. Both yield consistent high-redshift constraints, whereas the \texttt{srollv2} likelihood tightens the lower limit on the low-$z$ contribution, resulting on a tighter lower limit and higher mean for the total $\tau$. 
}
\label{fig:plot_taugtz_2015_vs_2018_simallEE_vs_2018_srollv2}
\end{figure}

@@

In conclusion, we used constraints on reionization principal components with respect to the large-scale $EE$ data to produce a fast and accurate effective likelihood for the final release of Planck 2018 data. In lieu of the official $\texttt{simallEE}$ likelihood for low-$\ell$ $EE$ we used the independent later release of the $\texttt{srollv2}$ likelihood in 2019 which has better foreground cleaning allowing for slightly improved parameter constraints. Our effective likelihood code package is called $\textsc{Relike}$ [to be changed] and is available publicly on Github. This code is beneficial for the community for the purpose of testing \textit{any} global ionization history between $6 < z < 30$ with the Planck data without having to modify CAMB which can be very cumbersome to do. The convergence time to reach the Gilman-Rubin criteria $R_{-1} < 0.01$ is about 300 times less with the Relike code than with a traditional MCMC sampling of the Planck likelihood for the one-parameter canonical tanh model.

>>>>

\begin{figure}[ht]
\includegraphics[width=0.45\textwidth]{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2.pdf}
\caption{Comparison of the 68\% and 95\% C.L. constraints on the cumulative optical depth with different Planck likelihoods. \textit{Top}: The Planck 2015 vs Planck 2018 likelihoods, the latter including the official low-$\ell$ $EE$ likelihood $\texttt{simall\_EE}$. \textit{Bottom}: The Planck 2018 likelihoods using $\texttt{simall\_EE}$ for the low-$\ell$ $EE$ likelihood vs \texttt{srollv2} which has better foreground cleaning. Both yield consistent high-redshift constraints, whereas the \texttt{srollv2} likelihood tightens the lower limit on the low-$z$ contribution, resulting on a tighter lower limit and higher mean for the total $\tau$. 
}
\label{fig:plot_taugtz_2015_vs_2018_simallEE_vs_2018_srollv2}
\end{figure}

@@

In conclusion, we used constraints on reionization principal components with respect to the large-scale $EE$ data to produce a fast and accurate effective likelihood for the final release of Planck 2018 data. In lieu of the official $\texttt{simall\_EE}$ likelihood for low-$\ell$ $EE$ we used the independent later release of the $\texttt{srollv2}$ likelihood in 2019 which has better foreground cleaning allowing for slightly improved parameter constraints. Our effective likelihood code package is called $\textsc{Relike}$ [to be changed] and is available publicly on Github. This code is beneficial for the community for the purpose of testing \textit{any} global ionization history between $6 < z < 30$ with the Planck data without having to modify CAMB which can be very cumbersome to do. The convergence time to reach the Gilman-Rubin criteria $R_{-1} < 0.01$ is about 300 times less with the Relike code than with a traditional MCMC sampling of the Planck likelihood for the one-parameter canonical tanh model.
